---
title: "RL simulation and parameter recovery with Stan II: More data"
---

**Can we get better group level parameter estimates if we had more data?**

```{r}
source('/Users/zeynepenkavi/Dropbox/RangelLab/BayesianStats/helpers/demo_QlearningSimulation.R')
```

## More trials per subject

### Example 1: Nice true parameter values

Increasing number of trials from 150 to 450.

```{r}
sim_data = demo_QlearningSimulation(alpha = .65, beta = 2.5, numBlocks = 18)
```

True learning rate `alpha`

```{r}
true_alpha = VBA_sigmoid(sim_data$simulation$evolution, inverse=FALSE)
true_alpha
```

True inverse temperature `beta`

```{r}
true_beta = exp(sim_data$simulation$observation)
true_beta
```

Use Stan to get parameters that generated simulated data.

```{r}
TN = length(sim_data$choice)
choice = sim_data$choices #adding 1 bc categorical_logit has support over [1,2]
outcome = sim_data$feedback
m_data = list(T = TN, choice = choice, outcome = outcome)
```

```{r}
m = stan_model('stanModels/QLearning_singleSubject.stan')
```

```{r}
fit_nuts = sampling(m, iter=1000, chains=4, data=m_data)
```

```{r}
data.frame(extract(fit_nuts, c("alpha", "beta"))) %>%
  mutate(alg = "NUTS") %>%
  gather(key, value, -alg) %>%
  group_by(key) %>%
  mutate(med_est = median(value),
         true_val = ifelse(key == "alpha", true_alpha, true_beta)) %>%

  gather(est_type, estimate, -alg, -key, -value) %>%
  
  ggplot(aes(x = value, fill=alg))+
  geom_histogram(position="identity", alpha = .5)+
  facet_grid(alg~key, scales='free')+
  geom_vline(aes(xintercept=estimate, linetype = est_type)) +
  scale_linetype_manual(name="", values=c("med_est" = "solid", "true_val" = "dashed"),
                        labels = c("med_est" = "median estimate", "true_val" = "true value"))+
  scale_fill_manual(name="", values=c("NUTS" = "purple"))+
  theme(legend.position="bottom")+
  xlab("")+
  ylab("")
```

```{r}
data.frame(extract(fit_nuts, c('alpha', 'beta', 'log_lik'))) %>%
  mutate(alg = "NUTS",
         est_alpha = median(alpha),
         est_beta = median(beta)) %>%
  mutate(true_alpha = true_alpha, true_beta = true_beta)%>%
  ggplot()+
  geom_point(aes(x=alpha, y=beta, color=log_lik)) +
  geom_point(aes(x=true_alpha, y = true_beta), color="red", shape = 13, size=6)+
  geom_point(aes(x=est_alpha, y = est_beta), color =  "black", shape=4, size=6)+
  theme(legend.position = "bottom")+
  ylab("Inverse temperature")+
  xlab("Learning rate")
```

### Example 2: High learning rate lower beta

```{r}
sim_data = demo_QlearningSimulation(alpha = .9, beta = .7, numBlocks = 18)
```

True learning rate `alpha`

```{r}
true_alpha = VBA_sigmoid(sim_data$simulation$evolution, inverse=FALSE)
true_alpha
```

True inverse temperature `beta`

```{r}
true_beta = exp(sim_data$simulation$observation)
true_beta
```

Use Stan to get parameters that generated simulated data.

```{r}
TN = length(sim_data$choice)
choice = sim_data$choices #adding 1 bc categorical_logit has support over [1,2]
outcome = sim_data$feedback
m_data = list(T = TN, choice = choice, outcome = outcome)
```

```{r}
fit_nuts = sampling(m, iter=1000, chains=4, data=m_data)
```

Having more trials does not help much if true parameter values are extreme.

```{r}
data.frame(extract(fit_nuts, c("alpha", "beta"))) %>%
  mutate(alg = "NUTS") %>%
  gather(key, value, -alg) %>%
  group_by(key) %>%
  mutate(med_est = median(value),
         true_val = ifelse(key == "alpha", true_alpha, true_beta)) %>%

  gather(est_type, estimate, -alg, -key, -value) %>%
  
  ggplot(aes(x = value, fill=alg))+
  geom_histogram(position="identity", alpha = .5)+
  facet_grid(alg~key, scales='free')+
  geom_vline(aes(xintercept=estimate, linetype = est_type)) +
  scale_linetype_manual(name="", values=c("med_est" = "solid", "true_val" = "dashed"),
                        labels = c("med_est" = "median estimate", "true_val" = "true value"))+
  scale_fill_manual(name="", values=c("NUTS" = "purple"))+
  theme(legend.position="bottom")+
  xlab("")+
  ylab("")
```

```{r}
data.frame(extract(fit_nuts, c('alpha', 'beta', 'log_lik'))) %>%
  mutate(alg = "NUTS",
         est_alpha = median(alpha),
         est_beta = median(beta)) %>%
  mutate(true_alpha = true_alpha, true_beta = true_beta)%>%
  ggplot()+
  geom_point(aes(x=alpha, y=beta, color=log_lik)) +
  geom_point(aes(x=true_alpha, y = true_beta), color="red", shape = 13, size=6)+
  geom_point(aes(x=est_alpha, y = est_beta), color =  "black", shape=4, size=6)+
  theme(legend.position = "bottom")+
  ylab("Inverse temperature")+
  xlab("Learning rate")
```

## Larger sample (and hierarchical model)

Generate simulated data using a model with a single learning `alpha` and inverse temperature parameter `beta` for 50 subjects.

```{r}
N = 50
true_g_alpha = .65
true_g_beta = 2.5
sim_data = list()
for (i in 1:N){
  cur_alpha = rnorm(1, mean = true_g_alpha, sd=.1)
  cur_beta = rnorm(1, mean = true_g_beta, sd = .5)
  sim_data[[i]] = demo_QlearningSimulation(alpha = cur_alpha, beta = cur_beta)
}
```

### Distribution of subject parameters

```{r}
df = data.frame(alphas=rep(NA, N), betas=rep(NA, N))

for (i in 1:N){
  df$alphas[i]=  VBA_sigmoid(sim_data[[i]]$simulation$evolution, inverse=FALSE)
  df$betas[i]= exp(sim_data[[i]]$simulation$observation)
}
```

```{r}
df %>% 
  gather(key, value) %>%
  ggplot(aes(x=value))+
  geom_histogram()+
  facet_wrap(~key, scales='free')
```
### Model fit

```{r}
TN = length(sim_data[[1]]$choice)
choice = matrix(nrow=N, ncol = TN)
outcome = matrix(nrow=N, ncol = TN)

for(i in 1:N){
  choice[i,] = sim_data[[i]]$choice
  outcome[i,] = sim_data[[i]]$feedback
}

m_data = list(N = N, T = TN, choice = choice, outcome = outcome)
```

```{r}
m = stan_model('stanModels/QLearning_hierarchical.stan')
```

```{r}
fit_nuts = sampling(m, iter=1000, chains=4, data=m_data)
```

### Posteriors

#### Group parameters

```{r}
labels <- c(g_alpha = "group alpha", g_beta = "group beta")

data.frame(extract(fit_nuts, c("g_alpha", "g_beta"))) %>%
  mutate(alg = "NUTS") %>%
  gather(key, value, -alg) %>%
  group_by(key) %>%
  mutate(med_est = median(value),
         true_val = ifelse(key == "g_alpha", true_g_alpha, true_g_beta)) %>%

  gather(est_type, estimate, -alg, -key, -value) %>%
  
  ggplot(aes(x = value, fill=alg))+
  geom_histogram(position="identity", alpha = .5)+
  facet_grid(alg~key, scales='free', labeller=labeller(key=labels))+
  geom_vline(aes(xintercept=estimate, linetype = est_type)) +
  scale_linetype_manual(name="", values=c("med_est" = "solid", "true_val" = "dashed"),
                        labels = c("med_est" = "median estimate", "true_val" = "true value"))+
  scale_fill_manual(name="", values=c("NUTS" = "purple"))+
  theme(legend.position="bottom")+
  xlab("")+
  ylab("")
```

#### Subject parameters

**Alphas**

```{r}
for(i in 1:N){
  df$key[i] = paste("Subj ", i, sep="")
}
data.frame(extract(fit_nuts, c("alpha"))) %>%
  gather(key, value) %>%
  mutate(key = gsub("alpha.", "Subj ", key)) %>%
  full_join(df %>% select(alphas, key), by='key') %>%
  group_by(key) %>% 
  mutate(med_est = median(value)) %>%
  rename(true_val=alphas) %>%
  gather(est_type, estimate, -key, -value) %>%
  ggplot(aes(value)) +
  geom_histogram(alpha = .5, fill="purple")+
  geom_vline(aes(xintercept = estimate, linetype=est_type))+
  scale_linetype_manual(name="", values=c("med_est" = "solid", "true_val" = "dashed"),
                        labels = c("med_est" = "median estimate", "true_val" = "true value"))+
    theme(legend.position="bottom")+
  facet_wrap(~key)+
  xlab("")+
  ggtitle("Subject alphas")

```

**Betas**

```{r}
data.frame(extract(fit_nuts, c("beta"))) %>%
  gather(key, value) %>%
  mutate(key = gsub("beta.", "Subj ", key)) %>%
  full_join(df %>% select(betas, key), by='key') %>%
  group_by(key) %>% 
  mutate(med_est = median(value)) %>%
  rename(true_val=betas) %>%
  gather(est_type, estimate, -key, -value) %>%
  ggplot(aes(value)) +
  geom_histogram(alpha = .5, fill="purple")+
  geom_vline(aes(xintercept = estimate, linetype=est_type))+
  scale_linetype_manual(name="", values=c("med_est" = "solid", "true_val" = "dashed"),
                        labels = c("med_est" = "median estimate", "true_val" = "true value"))+
  theme(legend.position="bottom")+
  facet_wrap(~key)+
  xlab("")+
  ggtitle("Subject betas")

```