---
title: "Neural mass model simulation and parameter recovery with Stan"
---

```{r}
library(tidygraph)
library(ggraph)
```

```{r}
source('~/Dropbox/RangelLab/BayesianStats/helpers/networkModel.R')
source('~/Dropbox/RangelLab/BayesianStats/helpers/make_stimtimes.R')
```

Suppose we have a three node network that looks like:

```{r}
edges = data.frame(from = c(1, 2, 2), to = c(2, 1, 3), weight = c(.4, .2, .3))
nodes = data.frame(id = c(1,2,3), label = c("1", "2", "3"))
```

```{r}
min_net = tbl_graph(nodes=nodes, edges=edges, directed=T)
min_net
```

```{r}
ggraph(min_net, layout="circle")+
  geom_edge_parallel(aes(width=weight, label=weight), 
                 alpha=.8,
                 arrow = arrow(length = unit(4, 'mm')),
                 end_cap = circle(5, 'mm'),
                 start_cap = circle(5, 'mm'),
                 label_dodge=unit(-4.0,"mm"),
                 label_push=unit(4,"mm"),
                 position="identity",angle_calc="along",force_flip=T)+
  scale_edge_width(range=c(.2,2))+
  geom_node_point(size=7)+
  geom_node_label(aes(label=label), 
                 repel=T)+ 
  theme_graph()+
  theme(legend.position = "none",
        plot.margin = margin(0, 1, .5, 1, "cm"))
```
Adjacency matrix for the above minimal network looks like:

```{r}
W<-matrix(0, 3, 3)
W[as.matrix(edges[,2:1])] <- edges$weight
W
```

```{r}
edges %>%
  rbind(data.frame(from = c(1,2,3,3,1,3), to = c(1,2,3,1,3,2), weight = c(0,0,0,0,0,0))) %>%
  ggplot(aes(x=from, y=to, fill=weight))+
  geom_tile()
```
Suppose we have a task that looks like:

```{r}
task = data.frame(stim = rep(c(0,0,0,1,0,0,0),3))
task$time = rep(1:nrow(task))
task %>%
  ggplot(aes(x=time, y=stim))+
  geom_line()+
  xlab("")+
  ylab("")+
  scale_x_continuous(breaks = seq(1,nrow(task),1))+
  scale_y_continuous(breaks = c(0,1))
```

The task stimulates *only* node 1 directly and propagates within the network based on the following Eq:

$$\frac{dx_i}{dt}\tau_i = -x_i(t) + s\phi\big(x_i(t)\big) + g\Bigg(\sum_{j\neq i}^{N} W_{ij}\phi\big(x_j(t)\big)\Bigg) + I_i(t)$$
Activity in each node at each time point would look like:

```{r}
cur_args_dict = list('dt'=.5,  
                 'g'=1, 
                 'noise'= NULL,
                 'noise_loc'= 0, 
                 'noise_scale'= 0,
                 's'=.7,
                 'stim_mag'=.5,
                 'taskdata'=NULL,
                 'tasktiming'=task$stim,
                 'tau'=1, 
                 'Tmax'=max(task$time),
                 'W'= W)

cur_args_dict$I = make_stimtimes(1, cur_args_dict)$stimtimes
```

```{r}
make_stimtimes(1, cur_args_dict)$tasktiming
```

```{r}
net_dat = networkModel(W, cur_args_dict)

data.frame(net_dat$Enodes) %>%
  mutate(node = c(1,2,3)) %>%
  gather(sampling, value, -node) %>%
  mutate(sampling = gsub("X", "", sampling),
         sampling = as.numeric(sampling),
         node = as.factor(node)) %>%
  ggplot(aes(x=sampling, y=value, color = node))+
  geom_line()
```



------

Question re estimation methods: Are the posteriors unimodal in VI vs MCMC? Are we losing important information when making assumptions like the Laplace approximation?
