---
title: "ODE benchmarking"
output: html_notebook
---

```{r}
library(cmdstanr)
library(posterior)
set.seed(1302589023)
```

# Generate fake data

```{r}
make_states = function(n=500, tmin = 1E-3, tmax=48, a0=10, ke=0.2, noise_mean = 0, noise_scale=.1){
  ts = seq(tmin, tmax, length=n)
  noise = rnorm(n, noise_mean, noise_scale)
  a = a0 * exp(-ke*ts) + noise
  out = data.frame(ts=ts, a=a, noise=noise)
  return(out)
}
```

View fake data

```{r}
dat = make_states()
dat %>%
  ggplot(aes(ts, a))+
  geom_point()
```
Set up data for Stan models

```{r}
mod_dat = list(N = length(dat$a),
                ts = dat$ts,
                a = matrix(dat$a, ncol=1))
```

# Fit models

Fit analytical model

$a(t) = a_0e^(k_et)$

```{r}
mod1 = cmdstan_model('./stanModels/odeTest1.stan')
```

```{r}
fit1 = mod1$sample(data = mod_dat)
fit1$save_object(file = './stanModels/fit_odeTest1.RDS')
```

Fit ODE model

$\frac{da}{dt} = -k_e * a(t)$

```{r}
mod2 = cmdstan_model('./stanModels/odeTest2.stan')
```

```{r}
fit2 = mod2$sample(data = mod_dat)
fit2$save_object(file = './stanModels/fit_odeTest2.RDS')
```

# Examine output

## Analytical solution

```{r}
fit1
```

## ODE solver

```{r}
fit2
```

## Posterior predictive check comparison

```{r}
ppc_df1 = fit1$summary("a_gen") %>%
  select(variable, mean)

ppc_df2 = fit2$summary("a_gen") %>%
  select(variable, mean)
```

```{r}
ppc_df1 %>%
  left_join(ppc_df2, by = "variable") %>%
  mutate(true=dat$a) %>%
  select(-variable) %>%
  rename(analytical=mean.x, ode_solver=mean.y) %>%
  gather(key, value, -true) %>%
  ggplot(aes(true, value))+
  geom_point()+
  facet_wrap(~key)+
  geom_abline(aes(intercept=0, slope=1))
```

# Model run time

## Analytical solution

```{r}
fit1$time()$total
```

## ODE solver

```{r}
fit2$time()$total
```

# Examine chains

Difference between `.draw` and `.iteration`: the former ranges from 1-4000 with 1000 values for each chain sequentially whereas the latter ranges from 1-1000 for each chain.

```{r}
as_draws_df(fit1$draws()) %>%
  select(ke, .iteration, .chain) %>%
  mutate(key='analytical') %>%
  rbind(as_draws_df(fit2$draws()) %>%
          select(ke, .iteration, .chain) %>%
          mutate(key='ode_solver')) %>%
  ggplot(aes(.iteration, ke))+
  geom_line()+
  facet_grid(.chain~key)+
  xlab("iteration")
```