---
title: "Acticity spread and eGLM success in larger network"
---

```{r}
library(tidygraph)
library(ggraph)
library(gridExtra)
```

```{r}
helpers_path = '~/Dropbox/RangelLab/NetworkGLM/helpers/r_helpers/'
source(paste0(helpers_path,'networkModel.R'))
source(paste0(helpers_path,'make_stimtimes.R'))
source(paste0(helpers_path,'generateStructuralNetwork.R'))
source(paste0(helpers_path,'generateSynapticNetwork.R'))
source(paste0(helpers_path,'run_ucr_glm.R'))
source(paste0(helpers_path,'run_ext_glm.R'))
source(paste0(helpers_path,'check_net_resids.R'))
source(paste0(helpers_path,'plot_adj_mat.R'))
source('~/Dropbox/PoldrackLab/SRO_Retest_Analyses/code/helper_functions/g_legend.R')
```

## Larger network

Does the extended GLM framework work for a larger network with a topdown block task?

```{r}
big_W_args_dict = default_args_dict

big_G = generateStructuralNetwork(big_W_args_dict)
big_W_args_dict$W = generateSynapticNetwork(big_G$W, showplot=big_W_args_dict$showplot)$G
big_W_args_dict$stim_nodes = c(1:10)

big_W_args_dict$tasktiming = c(rep(0, 99), rep(1, 100), rep(0, 301))

big_W_args_dict$I = make_stimtimes(big_W_args_dict$stim_nodes, big_W_args_dict)$stimtimes
big_W_net_dat = networkModel(big_W_args_dict$W, big_W_args_dict)
big_W_net_dat_debug = networkModel(big_W_args_dict$W, big_W_args_dict, debug=T)
```

Suppose we have a network of 105 nodes with 3 communities of 35 nodes each. This network has one hub and two local communities. The hub community is more likely to be connected to the local community nodes, while local communities only have high connectivity within their own networks.

```{r}
plot_adj_mat(big_W_args_dict$W)$p
```

First check network structure. Confirm that the number of total connections to stim nodes and the sum of weights from stim nodes is larger for non-stim hub nodes than they are for non stim local nodes.

```{r}
#from in columns; to in rows
dist_mat = igraph::distances(igraph::graph_from_adjacency_matrix(big_W_args_dict$W, weighted=T), algorithm = 'johnson')

tmp = data.frame(sum_stim_node_cons = rowSums(big_G$W[11:105, 1:10]), 
           sum_stim_node_weights = rowSums(big_W_args_dict$W[11:105, 1:10]),
           mean_distance = rowMeans(dist_mat[11:105 , 1:10])) %>%
  mutate(node = 1:n(),
         node = node+10,
         comm = ifelse(node<36, 1, ifelse(node>35 & node<71, 2, 3)))


tmp %>%
  gather(key, value, -node, -comm) %>%
  ggplot(aes(node, value))+
  geom_line()+
  geom_vline(aes(xintercept = 35), linetype="dashed")+
  geom_vline(aes(xintercept = 70), linetype="dashed")+
  facet_grid(key~., scales='free_y', switch = 'both')+
  scale_y_continuous(position="right")+
  ylab("")
```

The block task comes on first after 100 time points and then after every 400 time points for 100 time points. It is top-down, meaning it only stimulates some (10) nodes in the hub network directly.

```{r}
data.frame(stim = make_stimtimes(big_W_args_dict$stim_nodes, big_W_args_dict)$tasktiming) %>%
  mutate(timepoint = 1:n())%>%
  ggplot(aes(timepoint, stim)) +
  geom_line()
```

Activity propagates in this network based on the following self and network stimulation parameters

```{r}
cat(paste0("Self stimulation parameter s is: ", big_W_args_dict$s))
cat("\n")
cat(paste0("Network stimulation parameter g is: ", big_W_args_dict$g))
```

What does the activity look like in the network?

```{r}
data.frame(big_W_net_dat) %>%
  mutate(node = 1:n())  %>%
  gather(time, value, -node) %>%
  mutate(time = gsub("X", "", time),
         time = as.numeric(time),
         comm = ifelse(node<11, "1_stim", ifelse(node>10 & node<36, "1_nonstim", ifelse(node>35 & node<71, "2_nonstim", "3_nonstim")))) %>%
  ggplot(aes(time, value, col=comm))+
  geom_point()+
  theme(legend.title = element_blank())
```

```{r}
print(paste0("Stim node is: ", big_W_args_dict$stim_nodes))
cat("\n")
cat("Uncorrected estimates")
cat("\n")
round(run_ucr_glm(big_W_net_dat_debug, task_reg = big_W_args_dict$I[1,])$ucr_task_betas, 4)
cat("\n")
cat("Corrected estimates with network activity")
cat("\n")
round(run_ext_glm(big_W_net_dat_debug, big_W_args_dict, task_reg = big_W_args_dict$I[1,])$ext_task_betas, 4)
cat("\n")
cat("Corrected estimates without network activity")
cat("\n")
round(run_ext_glm(big_W_net_dat_debug, big_W_args_dict, task_reg = big_W_args_dict$I[1,], inc_net_act = FALSE)$ext_task_betas, 4)
```

Plotting as I did before to visually detect if there is any "bleeding" to non-stim hub nodes. 

```{r}
data.frame(ucr_betas = run_ucr_glm(big_W_net_dat_debug, task_reg = big_W_args_dict$I[1,])$ucr_task_betas,
           ext_betas = run_ext_glm(big_W_net_dat_debug, big_W_args_dict, task_reg = big_W_args_dict$I[1,])$ext_task_betas) %>%
  mutate(node = 1:n()) %>%
  gather(key, beta, -node) %>%
  ggplot(aes(node, beta, col=key))+
  geom_line()+
  theme(legend.title = element_blank())+
  geom_vline(aes(xintercept = 35), linetype="dashed")+
  geom_vline(aes(xintercept = 70), linetype="dashed")
```

**WHERE HAS THE BLEEDING GONE?**

More generally when does activity turn off?

```{r}
sweep_vals = c(seq(0,1.5,.1), seq(1, 5, .5))

sweep_W_args_dict = default_args_dict
sweep_G = generateStructuralNetwork(sweep_W_args_dict)
sweep_W_args_dict$W = generateSynapticNetwork(sweep_G$W, showplot=sweep_W_args_dict$showplot)$G
sweep_W_args_dict$stim_nodes = c(1:10)
sweep_W_args_dict$tasktiming = c(rep(0, 99), rep(1, 100), rep(0, 301))
sweep_W_args_dict$I = make_stimtimes(sweep_W_args_dict$stim_nodes, sweep_W_args_dict)$stimtimes

out = data.frame(s=rep(NA,length(sweep_vals)^2), g=rep(NA,length(sweep_vals)^2),
                 mean_stimnodes_end=rep(NA,length(sweep_vals)^2), mean_nonstimnodes_end=rep(NA,length(sweep_vals)^2),
                 mean_nonstim_hub_dev=rep(NA,length(sweep_vals)^2), mean_nonstim_local_dev=rep(NA,length(sweep_vals)^2))
cur_row=1
for(i in 1:length(sweep_vals)){
  for(j in 1:length(sweep_vals)){
    cur_s = sweep_vals[i]
    cur_g = sweep_vals[j]
    sweep_W_args_dict$s = cur_s
    sweep_W_args_dict$g = cur_g
    sweep_W_net_dat = networkModel(sweep_W_args_dict$W, sweep_W_args_dict)
    tmp = sweep_W_net_dat[,dim(sweep_W_net_dat)[2]]
    cur_stimend = mean(tmp[1:10])
    cur_nonstimend = mean(tmp[11:length(tmp)])
    
    ucr_betas = run_ucr_glm(sweep_W_net_dat, task_reg = sweep_W_args_dict$I[1,])$ucr_task_betas
    mean_nonstim_hub_dev = mean(ucr_betas[1:10]) - mean(ucr_betas[11:35])
    mean_nonstim_local_dev= mean(ucr_betas[1:10]) - mean(ucr_betas[36:105])
    
    out[cur_row,] = c(cur_s, cur_g, cur_stimend, cur_nonstimend, mean_nonstim_hub_dev, mean_nonstim_local_dev)
    cur_row = cur_row+1
  }
}
```

```{r}
out %>%
  select(-mean_nonstim_hub_dev, -mean_nonstim_local_dev) %>%
  mutate(s=as.factor(s),
         g=as.factor(g))%>%
  gather(key, value, -s, -g) %>%
  ggplot(aes(s, g, fill=value))+
  geom_tile()+
  facet_wrap(~key)+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle('Activity level in last time point')
```

```{r}
out %>%
  select(-mean_stimnodes_end, -mean_nonstimnodes_end) %>%
    filter(g>0) %>%
  mutate(s=as.factor(s),
         g=as.factor(g))%>%
  gather(key, value, -s, -g) %>%
  ggplot(aes(s, g, fill=value))+
  geom_tile()+
  facet_wrap(~key)+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle("Mean deviance of task betas from stim nodes' task betas")
```