---
title: "Neural mass model simulation and external stimulation parameter recovery with Stan"
---

```{r}
library(tidygraph)
library(ggraph)
library(gridExtra)
library(cmdstanr)
```

```{r}
helpers_path = '~/Dropbox/RangelLab/NetworkGLM/helpers/r_helpers/'
source(paste0(helpers_path,'networkModel.R'))
source(paste0(helpers_path,'make_stimtimes.R'))
source(paste0(helpers_path,'plot_adj_mat.R'))
```

# Developing intuitions for activity propogation within a network

Suppose we have a three node network that looks like:

```{r}
edges = data.frame(from = c(1, 2, 2), to = c(2, 1, 3), weight = c(.4, .2, .3))
nodes = data.frame(id = c(1,2,3), label = c("1", "2", "3"))
min_net = tbl_graph(nodes=nodes, edges=edges, directed=T)
```

```{r}
ggraph(min_net, layout="circle")+
  geom_edge_parallel(aes(width=weight, label=weight), 
                 alpha=.8,
                 arrow = arrow(length = unit(4, 'mm')),
                 end_cap = circle(5, 'mm'),
                 start_cap = circle(5, 'mm'),
                 label_dodge=unit(-4.0,"mm"),
                 label_push=unit(4,"mm"),
                 position="identity",angle_calc="along",force_flip=T)+
  scale_edge_width(range=c(.2,2))+
  geom_node_point(size=7)+
  geom_node_label(aes(label=label), 
                 repel=T)+ 
  theme_graph()+
  theme(legend.position = "none",
        plot.margin = margin(0, 1, .5, 1, "cm"))
```

Adjacency matrix for the above minimal network looks like:

```{r}
W<-matrix(0, 3, 3)
W[as.matrix(edges[,2:1])] <- edges$weight
W
```

```{r}
plot_adj_mat(W)$p
```

Suppose we have resting state data (i.e. no task).

Activity in each node at each time point would look like:

```{r}
cur_args_dict = list('dt'=.5,  
                 'noise'= TRUE,
                 'noise_loc'= 0, 
                 'noise_scale'= 0.1,
                 's'=.3,
                 'g'=.7,
                 'taskdata'=NULL,
                 'tau'=1, 
                 'Tmax'=100,
                 'W'= W)

net_dat = networkModel(W, cur_args_dict)

data.frame(net_dat) %>%
  mutate(node = c(1,2,3)) %>%
  gather(sampling, value, -node) %>%
  mutate(sampling = gsub("X", "", sampling),
         sampling = as.numeric(sampling),
         node = as.factor(node)) %>%
  ggplot(aes(x=sampling, y=value, color = node))+
  geom_line()
```

# Bayesian parameter recovery

```{r}
cat(paste0("True s is: ", cur_args_dict$s))
cat('\n')
cat(paste0("True g is: ", cur_args_dict$g))
cat('\n')
cat(paste0("True tau is: ", cur_args_dict$tau))
cat('\n')
cat(paste0("True sigma is: ", cur_args_dict$noise_scale))
```

```{r}
nmm_data = list(N_TS = dim(net_dat)[2],
                N = dim(net_dat)[1],
                ts = 1:dim(net_dat)[2],
                y_init = net_dat[,1],
                y = t(net_dat),
                W = cur_args_dict$W,
                g = cur_args_dict$g,
                tau = cur_args_dict$tau)
```

### Model without spont act in ODE

```{r}
mod_orig = cmdstan_model("stanModels/nmm_ode_orig.stan")
```

```{r}
fit_orig = mod_orig$variational(data=nmm_data)
```

```{r}
fit_orig
```

```{r}
posterior::as_draws_df(fit_orig$draws()) %>%
  select(s, sigma) %>%
  gather(key, value) %>%
  mutate(type="posterior") %>%
  ggplot()+
  geom_histogram(aes(value))+
  facet_wrap(~key, scales='free')
```

Posterior predictive checks

```{r}
ppc_df = fit_orig$summary("y_gen")
```

```{r}
ppc_df = ppc_df %>%
  select(variable, mean) %>%
  separate(variable, sep = "\\[", into=c("var", "tmp")) %>%
  separate(tmp, sep = ",", into=c("time", "node")) %>%
  mutate(node = gsub("]", "", node),
         time = as.numeric(time)) 
```

Estimated data has a lot less variability.

```{r}
ppc_df %>%
  mutate(key = "Est data") %>%
  select(-var) %>%
  rbind(data.frame(net_dat) %>%
          mutate(node = c(1,2,3)) %>%
          gather(time, mean, -node) %>%
          mutate(time = gsub("X", "", time),
                 time = as.numeric(time),
                 node = as.factor(node),
                 key = "True data")) %>%
  ggplot(aes(time, mean))+
  geom_line(aes(color=node))+
  facet_grid(node~key)+
  theme(legend.position = "none")
```

The model is estimating the SD of the timeseries!

```{r}
data.frame(net_dat) %>%
  mutate(node = c(1,2,3)) %>%
  gather(time, value, -node) %>%
  mutate(time = gsub("X", "", time),
         time = as.numeric(time),
         node = as.factor(node),
         key = "True data") %>%
  group_by(node) %>% 
  summarise(mean_ts = mean(value),
            sd_ts = sd(value))
```

### Model WITH spont act in ODE

```{r}
mod = cmdstan_model("stanModels/recover_s.stan")
```

```{r}
fit = mod$variational(data=nmm_data)
```

```{r}
fit
```

```{r}
posterior::as_draws_df(fit$draws()) %>%
  select(s, sigma, m_err) %>%
  gather(key, value) %>%
  mutate(key = factor(key, levels =c("s", "sigma", "m_err"))) %>%
  ggplot()+
  geom_histogram(aes(value), bins=20)+
  facet_wrap(~key, scales='free')+
  xlab("")
```

Posterior predictive checks

```{r}
ppc_df = fit$summary("y_gen")
```

```{r}
ppc_df = ppc_df %>%
  select(variable, mean) %>%
  separate(variable, sep = "\\[", into=c("var", "tmp")) %>%
  separate(tmp, sep = ",", into=c("time", "node")) %>%
  mutate(node = gsub("]", "", node),
         time = as.numeric(time)) 
```

Estimated data now has more similar variability.

```{r}
ppc_df %>%
  mutate(key = "Est data") %>%
  select(-var) %>%
  rbind(data.frame(net_dat) %>%
          mutate(node = c(1,2,3)) %>%
          gather(time, mean, -node) %>%
          mutate(time = gsub("X", "", time),
                 time = as.numeric(time),
                 node = as.factor(node),
                 key = "True data")) %>%
  ggplot(aes(time, mean))+
  geom_line(aes(color=node))+
  facet_grid(node~key)+
  theme(legend.position = "none")
```

------

MCMC is still very slow. 

Read [this](https://github.com/generable/stan-ode-workshop/blob/master/ODEs-in-Stan.pdf) more carefully for parallelization later 

```{r eval=FALSE}
fit_mcmc <- mod$sample(
  data = nmm_data,
  seed = 123
)
```


------

Question re estimation methods: Are the posteriors unimodal in VI vs MCMC? Are we losing important information when making assumptions like the Laplace approximation?
